---
title: "HW"
author: "Matthew L. Anderson"
date: "December 7th 2018"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 1
## a
H0: There are no differences between the mean infant birthweights in any of the three racial groups
Ha: There is a difference between the mean infant birthweights in at least 1 of the three racial groups

In order to ASSESS this claim, we will conduct a 1-way ANOVA. First we will provide a visual assesment of the data and summary statistics



```{r cars, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(car)
BMI <- read_csv("~/VCU/Biomedical_Data_1/HW5/BMI.csv", col_names = TRUE)

BMI<-as.data.frame(BMI)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
boxplot(BMI$Birthweight ~ BMI$`Maternal Racial Group`,
        main="Birthweight by Racial Group",
        ylab="Weight",
        ylim=c(1000, 5500),
        xlab="Racial Group",
        col="grey",
        boxwex=0.3)

#Summary Statistics of the data
summary(BMI$Birthweight)
hist(BMI$Birthweight)
```

Table 1 shows the summary statisitcs(mean, sd, se, lower, upper) for each group.
BOxplot 

1-way ANOVA Model
```{r}
#will perform one way anova to see difference in means

# 1 way ANOVA Model 
model_BMI <- lm(BMI$Birthweight ~  factor(BMI$`Maternal Racial Group`), data = BMI)
anova(model_BMI)
summary(model_BMI)
```
Table 1. Summary statistics 

Assumptions of ANOVA Model 
```{r pressure, echo=FALSE}
#assumptions of normality by checking QQ plot of residuals 
plot(model_BMI, which = 1) 
qqnorm(resid(model_BMI)) 
qqline(resid(model_BMI))
```



```{r}
#Test for normality of response variable
shapiro.test(resid(model_BMI))

#Test for homogeniety of varaince 
leveneTest(Birthweight ~ factor(`Maternal Racial Group`), center = mean, data = BMI)
```


Using a 1-way ANOVA no differences in infant birthweight were found between the groups (F=1.3917, P-value=0.2815). The data do not depart from a normal distribution using the qq plot of the residulas and a shapiro test(p-value = 0.9424). 

## b
```{r pressur, echo=FALSE}
#Appears linear, some heteroskedacitity
plot(BMI$Birthweight,BMI$`Maternal BMI`)
```

```{r}
corrs<-cor(BMI$Birthweight, BMI$`Maternal BMI`, method = "spearman")

full_model<-lm(BMI$Birthweight~ BMI$`Maternal BMI`)

summary(full_model)
```

## c
```{r pressureel, echo=FALSE}
#assesment of normality of the response variable birthweight
#plot qq plot of residuals of response variable 
plot(full_model, which=1)
plot(full_model, which=2)

resid(full_model) #List of residuals
plot(density(resid(full_model))) #A density plot
qqnorm(resid(full_model)) # A quantile normal plot - good for checking normality
qqline(resid(full_model))

```
Figure 2. QQ Plot of the Residuals 

```{r}
#Test for normality of response variable
shapiro.test(resid(full_model))
```


#d
```{r}
ancova_1 <- aov(BMI$Birthweight ~ BMI$`Maternal BMI` + factor(`Maternal Racial Group`), data = BMI)
# NOTE: covariate goes first!! NOTE: there is
# no interaction
summary(ancova_1)

TukeyHSD(ancova_1)
```





#Question 2
## a
H0: There are no differences between the mean infant birthweights in any of the three racial groups
Ha: There is a difference between the mean infant birthweights in at least 1 of the three racial groups

In order to ASSESS this claim, we will conduct a 1-way ANOVA. First we will provide a visual assesment of the data and summary statistics


```{r carsq, echo=FALSE, message=FALSE, warning=FALSE}
HITS <- read_csv("~/VCU/Biomedical_Data_1/HW5/HITS.csv", col_names = TRUE)

HITS<-as.data.frame(HITS)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#first I want to just do a visual inspection of the data

boxplot(HITS$HITS_Score ~ HITS$Group,
        main="HITS Score by IPV Group",
        ylab="HITS Score ",
        ylim=c(0, 25),
        xlab="IPV Group",
        col="grey",
        boxwex=0.3)

#Summary Statistics of the data
summary(HITS$HITS_Score)
hist(HITS$HITS_Score)
```

Table 1 shows the summary statisitcs(mean, sd, se, lower, upper) for each group.
BOxplot 

1-way ANOVA Model
```{r}
# 1 way ANOVA Model 
model_HITS <- lm(HITS$HITS_Score ~  factor(HITS$Group), data = HITS)
anova(model_HITS)
summary(model_HITS)

```
Table 1. Summary statistics 

Assumptions of ANOVA Model 
```{r pressurew, echo=FALSE}
#assumptions of normality by checking QQ plot of residuals 
plot(model_HITS, which = 1) 
qqnorm(resid(model_HITS)) 
qqline(resid(model_HITS))

```



```{r}
#Test for normality of response variable
shapiro.test(resid(model_HITS))

#Test for homogeniety of varaince 
leveneTest(HITS_Score ~ factor(Group), center = mean, data = HITS)

```


Using a 1-way ANOVA no differences in infant birthweight were found between the groups (F=1.3917, P-value=0.2815). The data do not depart from a normal distribution using the qq plot of the residulas and a shapiro test(p-value = 0.9424). 

## b
```{r pressuree, echo=FALSE}
#Appears linear, some heteroskedacitity
plot(HITS$HITS_Score,HITS$Age)
```

```{r}
corrs<-cor(HITS$HITS_Score, HITS$Age, method = "pearson")

full_model<-lm(HITS$HITS_Score~ HITS$Group)

summary(full_model)
corrs
```

## c
```{r pressureea, echo=FALSE}
#assesment of normality of the response variable birthweight
#plot qq plot of residuals of response variable 
plot(full_model, which=1)
plot(full_model, which=2)

resid(full_model) #List of residuals
plot(density(resid(full_model))) #A density plot
qqnorm(resid(full_model)) # A quantile normal plot - good for checking normality
qqline(resid(full_model))

```
Figure 2. QQ Plot of the Residuals 

```{r}
#Test for normality of response variable
shapiro.test(resid(full_model))
```


#d
```{r}

ancova_1 <- aov(HITS$HITS_Score ~ HITS$Age + factor(Group), data = HITS)
# NOTE: covariate goes first!! NOTE: there is
# no interaction
summary(ancova_1)

#TukeyHSD(ancova_1)

```









