---
title: The Effect of Recombinant Human Growth Hormone (rhGH) and Insulin-Like Growth
  Factor-I (rhIGF-I) on Immune Response in Rhesus Monkeys
author: "Matthew L. Anderson"
date: "November 29, 2018"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Methods
Rhesus monkeys (n=20) were assinged to either rhGH (n=5), rhIGF-I (n=6), both rhGH and rhIGF-I (n=6) or a control (n=4). The assay for the in vivo function of the immune system relied on the response to an immunization with teanus toxoid. Means, standard deviations, standard errors, and 95% confidence intervals for the mean values are presented in Table 1. The data were analyzed using a 1-way ANOVA and a 2-way ANOVA to compare the immune responses in each group. A boxplot was used for a visual inspection of the data. Assumptions of normality of the residuals was checked using a qq plot and shapiro test. Homogeniety of varaince was also checked using residual plot and Leven's test using the mean. The F statistics and P-values were calcualted for each test as well as a contrasts to see the effect of each group on the model.

#Results 

```{r cars, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(car)
monkeys <- read_csv("./monkeys.csv", col_names = TRUE)
monkeys<-as.data.frame(monkeys)
se <- function(x) sqrt(var(x)/length(x))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
mean_<-c(mean(monkeys[1:4,2]),mean(monkeys[5:10,2]),mean(monkeys[11:15,2]),mean(monkeys[16:20,2]))
sd_<-c(sd(monkeys[1:4,2]),sd(monkeys[5:10,2]),sd(monkeys[11:15,2]),sd(monkeys[16:20,2]))
sd_error<-c(se(monkeys[1:4,2]),se(monkeys[5:10,2]),se(monkeys[11:15,2]),se(monkeys[16:20,2]))
ci_lower<-c((10.25-qt(0.975,df=4-1)*.4856267),(10.80-qt(0.975,df=6-1)*.5790797),(11.25-qt(0.975,df=5-1)*.2549510),
            ((11.52-qt(0.975,df=5-1)*.2922328)))
ci_upper<-c((10.25+qt(0.975,df=4-1)*.4856267),(10.80+qt(0.975,df=6-1)*.5790797),(11.25+qt(0.975,df=5-1)*.2549510),
            ((11.52+qt(0.975,df=5-1)*.2922328)))
summary_stat<-as.table(cbind(mean_,sd_,sd_error,ci_lower,ci_upper))
rownames(summary_stat) <- c("control", "rhigf", "rhgh", "rhigf and rhgh")
```

Table 1 shows the summary statisitcs(mean, sd, se, lower, upper) for each group.

```{r}
summary_stat
```
Table 1. Summary statistics 

```{r pressure, echo=FALSE}
#first I want to just do a visual inspection of the data

boxplot(monkeys$Response ~ monkeys$Group,
        main="Immune Function by Treatment",
        ylab="Immune Function",
        ylim=c(9, 14),
        xlab="Treatment",
        col="grey",
        boxwex=0.3)
```
Figure 1. Box Plot of Immune Function by Treatment


```{r}
# 1 way ANOVA Model 
model1 <- lm(monkeys$Response ~ monkeys$Group, data = monkeys)
anova(model1)
summary(model1)
```

Using a 1-way ANOVA no differences in mean Immune Function were found between the groups (F=1.3917, P-value=0.2815). The data do not depart from a normal distribution using the qq plot of the residulas and a shapiro test(p-value = 0.9424). 
```{r pressur, echo=FALSE}
#assumptions of normality by checking QQ plot of residuals 
plot(model1, which = 1) 
```
Figure 1. Residual Plot of Immune Function

```{r pressuree, echo=FALSE}
qqnorm(resid(model1)) 
qqline(resid(model1))
```
Figure 2. QQ Plot of the Residuals 

```{r}
#Test for normality of response variable
shapiro.test(resid(model1))
```

Using Leven's test for homogeniety of variance there is no evidence that the variances are unequal in each group.
```{r message=FALSE, warning=FALSE}
#Test for homogeniety of varaince 
leveneTest(Response ~ Group, center = mean, data = monkeys)
```


Using a 2-way ANOVA no differences in mean Immune Function were found between the groups (F=1.3917, P-value=0.2815). The same assumptions of normality and equal variance are meet.
```{r}
# 2 way ANOVA Model

#here I am creating two variables that assing a 0 or 1 to either containing rhigf or rhgh
monkeys$rhgh  <- ifelse(monkeys$Group=="Control"|monkeys$Group=="rhIGF-I",0,1)  
monkeys$rhigf<- ifelse(monkeys$Group=="Control"|monkeys$Group=="rhGH",0,1)   

#running the model with the interaction term
model2<-lm(Response ~ rhigf + rhgh + rhigf*rhgh, data = monkeys)
anova(model2)
summary(model2)
```

To show the individual effect of each group a contrast varaible was created. No significant effects were found. 
```{r}
#contrast and summary
Anova(model2, contrasts = list(topic=contr.sum, sys=contr.sum, type=3))
```










